---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(tidyverse)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300, 
  dev = "ragg_png"
)
```

# fitness

<!-- badges: start -->
<!-- badges: end -->

A dumping ground for some of my fitness data.

## exercise-minutes

I exported all of my Apple Health data. The export process provides a .zip file.
I opened the `export.xml` file in the .zip archive and extracted my daily
exercise minutes (the green ring) activity using the following:

```{r}
#| eval: false
library(tidyverse)

# Get all nodes in the document with summary
nodes_summary <- xml2::read_xml("C:/Users/trist/Desktop/export.xml") |> 
  xml2::xml_find_all(".//ActivitySummary") 

# Get the attributes in each node and create a dataframe from them
nodes_summary |> 
  xml2::xml_attrs() |>
  lapply(as.list) |> 
  lapply(as.data.frame) |> 
  bind_rows() |> 
  as_tibble() |> 
  # I got my first Apple Watch around Christmas 2018
  filter(dateComponents > "2018-12-19") |> 
  select(
    date = dateComponents,
    exercise_minutes = appleExerciseTime
  ) |> 
  type_convert(
    col_types = cols(
      date = col_date(format = ""),
      exercise_minutes = col_double(),
    )
  ) |>
  write_csv("data/exercise-minutes.csv")
```

Now, I have my exercise minutes for each day.

```{r exercise-minutes}
library(tidyverse)

data_minutes <- "data/exercise-minutes.csv" |> 
  # I added entries with Excel and it mangled the dates
  read_csv(
    col_types = cols(
      date = col_date("%m/%d/%Y"), 
      exercise_minutes = "i")
  ) |> 
  # in case i want to aggregate by months within years
  mutate(
    year = lubridate::year(date),
    month = lubridate::month(date),
    year_month = paste0(year, month)
  ) |> 
  group_by(year_month) |> 
  mutate(
    year_month_start = min(date)
  ) |> 
  ungroup()

ggplot(data_minutes |> filter(year > 2018)) + 
  aes(x = date, y = exercise_minutes) + 
  geom_hline(
    yintercept = 30, 
    color = "darkgreen", 
    size = 1, 
    linetype = "dashed"
  ) +
  geom_point(alpha = .4) +
  facet_wrap("year", scales = "free") + 
  stat_smooth(method = "loess", formula = y ~ x)
```

What is my exercise debt (unmet goal minutes)?

```{r}
data_minutes |> 
  group_by(year) |> 
  summarise(
    n_days = n(),
    mean_exercise = mean(exercise_minutes),
    exercise = sum(exercise_minutes),
    debt = sum(30 - exercise_minutes)
  )

data_minutes |> 
  summarise(
    n_days = n(),
    mean_exercise = mean(exercise_minutes),
    exercise = sum(exercise_minutes),
    debt = sum(30 - exercise_minutes)
  )
```


## intervals

  - Run some distance as fast as you can (time it), stop and recover until
    heartrate is 140 bpm (time how long the recovery takes).
  - Do this for the following distances (m): 800, 800, 600, 600, 400, 400, 200.

```{r intervals, fig.width = 8, fig.height = 4}
library(tidyverse)

convert_min_sec_to_sec <- function(xs) {
  # ye olde splitten applyen combinen
  xs |> 
    strsplit(split = ":", fixed = TRUE) |> 
    lapply(function(x) {
      x <- as.numeric(x)
      60 * x[1] + x[2]
    }) |> 
    unlist(use.names = FALSE)
}

d <- "data/intervals.csv" |> 
  read_csv(
    col_types = cols(
      interval_number = "i",
      time = "c",
      distance = "c",
      date = col_date("%m/%d/%Y")
    )
  ) |> 
  mutate(
    time = convert_min_sec_to_sec(time)
  )

d_date_margins <- d |> 
  group_by(date, interval_type) |> 
  summarise(
    intervals = n(),
    total_time = sum(time),
    distance = sum(as.numeric(distance)),
    .groups = "drop"
  )
d_date_margins

d |> 
  filter(interval_type == "run") |> 
  ggplot() + 
    aes(
      x = (interval_number + 1) / 2, 
      y = (time / 60) / (as.numeric(distance) / 1000)
    ) +
    geom_line(aes(group = date, color = factor(date))) +
    geom_point(aes(color = factor(date))) +
    geom_label(
      aes(
        label = paste0(distance, " m"), 
        y = 3.5
      ), 
      hjust = .5,
      stat = "unique"
    ) +
    geom_label(
      aes(label = label, y = 3.5),
      data = data.frame(interval_number = 15, label = "(all)"),
      hjust = .5,
    ) +
    geom_point(
      aes(color = factor(date)), 
      data = d_date_margins |> 
        filter(interval_type == "run") |> 
        mutate(interval_number = 15) |> 
        rename(time = total_time),
    ) +
    labs(
      x = "running interval number", 
      y = "pace [min per km] (lower: better)", 
      color = "date"
    )

d |> 
  filter(interval_type == "run") |> 
  ggplot() + 
    aes(
      x = (interval_number + 1) / 2, 
      y = (time) / (as.numeric(distance) / 200)
    ) +
    geom_line(aes(group = date, color = factor(date))) +
    geom_point(aes(color = factor(date))) +
    geom_label(
      aes(
        label = paste0(distance, " m"), 
        y = 30
      ), 
      hjust = .5,
      stat = "unique"
    ) +
    geom_label(
      aes(label = label, y = 30),
      data = data.frame(interval_number = 15, label = "(all)"),
      hjust = .5,
    ) +
    geom_point(
      aes(color = factor(date)), 
      data = d_date_margins |> 
        filter(interval_type == "run") |> 
        mutate(interval_number = 15) |> 
        rename(time = total_time),
    ) +
    labs(
      x = "running interval number", 
      y = "seconds per 200 m (lower: better)", 
      color = "date"
    )

d |> 
  filter(interval_type == "recover") |> 
  ggplot() + 
    aes(
      x = (interval_number + 1) / 2, 
      y = time
    ) +
    geom_line(aes(group = date, color = factor(date))) +
    geom_point(aes(color = factor(date))) +
    geom_label(
      aes(
        label = paste0(distance, " m"), 
        y = 45
      ), 
      hjust = .5,
      stat = "unique"
    ) +
    labs(
      x = "running interval number", 
      y = "recovery duration [s] (lower: better)", 
      color = "date",
      caption = "I used a different workout app on 2022-08-16 🤷‍♀️"
    )
```

## bodyweight

In April, I took up a new hobby: Weekly boxing classes. This new activity
motivated me to be a little more active (closing my watch's daily exercise goal
more often or trying exercises that help for boxing conditioning). I started
measuring my bodyweight around this time. I haven't change my diet or anything
like that, and I don't have any weight goals besides getting back to my
pre-pandemic bodyweight (250 lbs).

```{r bodyweight, fig.width = 6, fig.height = 4}
data_weight <- "data/bodyweight.csv" |> 
  read_csv(col_types = cols(date = col_date("%m/%d/%Y"))) 

p <- data_weight |> 
  ggplot() + 
  aes(x = date, y = weight) + 
  ylim(250, 270) + 
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_point() +
  theme_grey(base_size = 16) +
  labs(y = "bodyweight [lb]") 

p

library(ggrepel)

p + 
  geom_vline(
    data = NULL,
    linetype = "dashed",
    xintercept = as.Date("2022-04-24"),
    size = 1
  ) +
  geom_text_repel(
    data = data.frame(
      date = as.Date("2022-04-25"),
      weight = 269
    ),
    nudge_y = .5,
    point.padding = 0.2,
    segment.curvature = 1e-20,
    xlim = c(as.Date("2022-04-30"), NA),
    label = "Started taking boxing classes in late April 🥊",
    size = 5
  ) +
  geom_text_repel(
    data = data.frame(
      date = as.Date("2022-04-26"),
      weight = 250
    ),
    nudge_y = 1,
    point.padding = 0.5,
    segment.curvature = 1e-20,
    xlim = c(as.Date("2022-04-30"), NA),
    label = "Pre-pandemic weight was around 250 lbs",
    size = 5
  )  + 
  geom_blank(
    aes(x = max(c(data_minutes$date, data_weight$date)), y = 260),
  )
  
  
p2 <- last_plot()

p3 <- ggplot(data_minutes |> filter(date > as.Date("2022-01-01"))) + 
  aes(x = date, y = exercise_minutes) + 
  geom_hline(
    yintercept = 30, 
    color = "darkgreen", 
    size = 1, 
    linetype = "dashed"
  ) +
  geom_point(alpha = .4) +
  stat_smooth(method = "loess", formula = y ~ x) +
  labs(y = "Exercise minutes ⌚") + 
  theme_grey(base_size = 16) + 
  geom_vline(
    data = NULL,
    linetype = "dashed",
    xintercept = as.Date("2022-04-24"),
    size = 1
  ) +
  geom_blank(
    aes(x = max(c(data_minutes$date, data_weight$date)), y = 30),
  )

```

Not happy with this, yet.

```{r bodyweight2, fig.width = 10, fig.height = 8}
library(patchwork)

plot_spacer() + 
  p3 + 
  inset_element(
    p2, 
    left = 0.4, 
    bottom = 1, 
    right = 1, 
    top = 1.8, 
    align_to = "full"
  ) + 
  plot_layout(nrow = 2)
```

