---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
library(tidyverse)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300, 
  dev = "ragg_png"
)
```

# fitness

<!-- badges: start -->
<!-- badges: end -->

A dumping ground for some of my fitness data.

## exercise-minutes

I exported all of my Apple Health data. The export process provides a .zip file.
I opened the `export.xml` file in the .zip archive and extracted my daily
exercise minutes (the green ring) activity using the following:

```{r}
#| eval: false
library(tidyverse)

# Get all nodes in the document with summary
nodes_summary <- xml2::read_xml("C:/Users/trist/Desktop/export.xml") |> 
  xml2::xml_find_all(".//ActivitySummary") 

# Get the attributes in each node and create a dataframe from them
nodes_summary |> 
  xml2::xml_attrs() |>
  lapply(as.list) |> 
  lapply(as.data.frame) |> 
  bind_rows() |> 
  as_tibble() |> 
  # I got my first Apple Watch around Christmas 2018
  filter(dateComponents > "2018-12-19") |> 
  select(
    date = dateComponents,
    exercise_minutes = appleExerciseTime
  ) |> 
  type_convert(
    col_types = cols(
      date = col_date(format = ""),
      exercise_minutes = col_double(),
    )
  ) |>
  write_csv("data/exercise-minutes.csv")
```

Now, I have my exercise minutes for each day.

I should disclose that I turn on an Apple workout on my watch when I mow the
lawn, shovel snow, or do something mildly active like going for a walk with my
family. 

```{r exercise-minutes, fig.height = 7, fig.width = 10}
library(tidyverse)

data_minutes <- "data/exercise-minutes.csv" |> 
  # I added entries with Excel and it mangled the dates
  read_csv(
    col_types = cols(
      date = col_date("%m/%d/%Y"), 
      exercise_minutes = "i")
  ) |> 
  # in case i want to aggregate by months within years
  mutate(
    year = lubridate::year(date),
    month = lubridate::month(date),
    year_month = paste0(year, month)
  ) |> 
  group_by(year_month) |> 
  mutate(
    year_month_start = min(date)
  ) |> 
  ungroup()

ggplot(data_minutes |> filter(year > 2018)) + 
  aes(x = date, y = exercise_minutes) + 
  geom_hline(
    yintercept = 30, 
    color = "#92E82A", 
    linewidth = 1, 
  ) +
  geom_point(
    data = function(x) filter(x, exercise_minutes < 125),
    alpha = .4
  ) +
  geom_point(
    aes(y = 125),
    data = function(x) filter(x, exercise_minutes >= 125),
    alpha = .4,
    color = "orange"
  ) +
  geom_blank(
    data = tibble::tibble(
      exercise_minutes = NA_real_,
      date = as.Date("2023-12-31"),
      year = 2023
    )
  ) +
  facet_wrap("year", scales = "free_x") + 
  stat_smooth(method = "loess", formula = y ~ x) +
  scale_x_date(date_labels = "%b", date_minor_breaks = "1 months") +
  scale_y_continuous(breaks = c(0, 30, 60, 90, 120)) +
  labs(
    x = "month",
    y = "exercise activity minutes ⌚",
    caption = "Outliers (125+ minutes) replaced with orange points."
  ) +
  theme_light(base_size = 16)
```

I think that the smoothing lines should be match up from facet to facet. Let's
figure out how to do that.

```{r exercise-minutes-2, fig.height = 7, fig.width = 10}
library(tidyverse)

gam <- mgcv::gam(
  exercise_minutes ~ s(as.numeric(date), bs = "cr"), 
  data = data_minutes
)
data_gam <- gratia::fitted_values(gam, data = data_minutes)

ggplot(data_gam |> filter(year > 2018)) + 
  aes(x = date, y = exercise_minutes) + 
  geom_hline(
    yintercept = 30, 
    color = "#92E82A", 
    linewidth = 1, 
  ) +
  geom_point(
    data = function(x) filter(x, exercise_minutes < 125),
    alpha = .4
  ) +
  geom_point(
    aes(y = 125),
    data = function(x) filter(x, exercise_minutes >= 125),
    alpha = .4,
    color = "orange"
  ) +
  geom_blank(
    data = tibble::tibble(
      exercise_minutes = NA_real_,
      date = as.Date("2023-12-31"),
      year = 2023
    )
  ) +
  geom_smooth(
    aes(y = fitted, ymin = lower, ymax = upper),
    stat = "identity"
  ) +
  facet_wrap("year", scales = "free_x") + 
  scale_x_date(date_labels = "%b", date_minor_breaks = "1 months") +
  scale_y_continuous(breaks = c(0, 30, 60, 90, 120)) +
  labs(
    x = "month",
    y = "exercise activity minutes ⌚",
    caption = "Outliers (125+ minutes) replaced with orange points."
  ) +
  theme_light(base_size = 16)
```




What is my exercise debt (unmet goal minutes)?

```{r}
data_minutes |> 
  group_by(year) |> 
  summarise(
    n_days = n(),
    mean_exercise = mean(exercise_minutes),
    exercise = sum(exercise_minutes),
    debt = sum(30 - exercise_minutes)
  )

data_minutes |> 
  summarise(
    n_days = n(),
    mean_exercise = mean(exercise_minutes),
    exercise = sum(exercise_minutes),
    debt = sum(30 - exercise_minutes)
  )
```

My stupid Apple Fitness said my September 2022 challenge is 64 minutes of
exercise activity per day on average, so let's make a query to check on that.

```{r}
data_minutes |> 
  filter(year == 2022) |> 
  group_by(year, month) |> 
  summarise(
    num_days = n(),
    sum_minutes = sum(exercise_minutes),
    mean_minutes = scales::label_number(.1)(mean(exercise_minutes)),
    prop_of_sept_22_goal = (sum(exercise_minutes) / (64 * 30)) |> round(2),
    .groups = "drop"
  ) |> 
  ungroup() |> 
  arrange(desc(year), desc(month)) |> 
  knitr::kable(align = "lrrrr")
```

## bodyweight

In April, I took up a new hobby: Weekly boxing classes. This new activity
motivated me to be a little more active (closing my watch's daily exercise goal
more often or trying exercises that help for boxing conditioning). I started
measuring my bodyweight around this time. I haven't change my diet or anything
like that, and I don't have any weight goals besides getting back to my
pre-pandemic bodyweight (250 lbs).

```{r bodyweight, fig.width = 6, fig.height = 4}
data_weight <- "data/bodyweight.csv" |> 
  read_csv(col_types = cols(date = col_date("%m/%d/%Y"))) 

p <- data_weight |> 
  ggplot() + 
  aes(x = date, y = weight) + 
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_point() +
  ylim(235, 270) + 
  labs(y = "bodyweight [lb]") +
  theme_grey(base_size = 16) +
  theme(
    panel.grid.minor.y = element_blank()
  )

p

library(ggrepel)

data_weight |> 
  ggplot() + 
  aes(x = date, y = weight) + 
  geom_hline(
    yintercept = 250,
    linewidth = 1.5,
    color = "grey60"
  ) +
  geom_vline(
    linetype = "dashed",
    xintercept = as.Date("2022-04-24"),
    linewidth = 1,
  ) +
  stat_smooth(method = "loess", formula = y ~ x) +
  geom_point() +
  ylim(235, 270) + 
  labs(y = "bodyweight [lb]") +
  theme_grey(base_size = 16) +
  theme(
    # panel.grid.minor.y = element_blank()
  ) +
  geom_text_repel(
    data = data.frame(
      date = as.Date("2022-04-25"),
      weight = 269
    ),
    nudge_y = .5,
    point.padding = 0.2,
    segment.curvature = 1e-20,
    xlim = c(as.Date("2022-04-30"), NA),
    label = "Started taking boxing classes in late April 🥊",
    size = 4.5
  ) +
  geom_text_repel(
    data = data.frame(
      date = as.Date("2022-04-26"),
      weight = 250
    ),
    nudge_y = 0,
    point.padding = 0.5,
    segment.curvature = 1e-20,
    xlim = c(as.Date("2022-04-30"), NA),
    label = "Pre-pandemic weight\naround 250 lbs",
    size = 4.5
  )  + 
  geom_blank(
    aes(x = max(c(data_minutes$date, data_weight$date)), y = 260),
  )
  
```

The last time I was on a months-long fitness kick, I did a Nike+ Run Club
training program for a half-marathon. In a one-month period (July 2015), I ran a
half-marathon distance three times. My weight at that time was 234 lbs, so
that's the benchmark for Tristan, The Cardio Machine.


## other notes

11/25/2022 - got the cross-rope "get lean" weighted jump rope set

12/5/2022 - increased my apple move goal from 860 to 940

01/10/2023 - decreased my apple move goal to 500 because of a family vacation 
and procedure. started bumping it back up in 02/2023.

02/20/2023 - started doing intermittent fasting on weekdays. basically, this 
kept me from eating peanut butter before bed or in the middle of the night.



## archived/iced things

### fast miles

In September 2022, I started jogging my kid to preschool in a running stroller.
School is just a little more than a mile away. After a few weeks, I decided to
compete with myself for how fast I could run the first mile split on these jogs.

```{r fast-miles}
convert_min_sec_to_sec <- function(xs) {
  # ye olde splitten applyen combinen
  xs |> 
    strsplit(split = ":", fixed = TRUE) |> 
    lapply(function(x) {
      x <- as.numeric(x)
      60 * x[1] + x[2]
    }) |> 
    unlist(use.names = FALSE)
}

data_miles <- "data/miles-to-school.csv" |> 
  # I added entries with Excel and it mangled the dates
  read_csv(
    col_types = cols(
      date = col_date("%m/%d/%Y"), 
      first_mile_split = "c",
      comment = "c",
    )
  ) |> 
  mutate(
    time = convert_min_sec_to_sec(first_mile_split)
  )

ggplot(data_miles) + 
  aes(x = date, y = time) +
  geom_point(size = 3) +
  scale_y_continuous(
    breaks = (14:22 / 2) * 60,
    labels = function(x) {
      sprintf("%02d:%02d", x %/% 60, x %% 60)
    }
  ) +
  labs(
    x = "date", 
    y = "pace [min per mi]"
  ) +
  theme_light(base_size = 16)
```


### intervals

  - Run some distance as fast as you can (time it), stop and recover until
    heartrate is 140 bpm (time how long the recovery takes).
  - Do this for the following distances (m): 800, 800, 600, 600, 400, 400, 200.

I injured my knee doing the 200 m sprint on 2022-09-06 and now I'm kind of
scared of this workout.

```{r intervals, fig.width = 8, fig.height = 4}
library(tidyverse)
d <- "data/intervals.csv" |> 
  read_csv(
    col_types = cols(
      interval_number = "i",
      time = "c",
      distance = "c",
      date = col_date("%m/%d/%Y")
    )
  ) |> 
  mutate(
    time = convert_min_sec_to_sec(time)
  )

d_date_margins <- d |> 
  group_by(date, interval_type) |> 
  summarise(
    intervals = n(),
    total_time = sum(time),
    distance = sum(as.numeric(distance)),
    .groups = "drop"
  )
d_date_margins

d |> 
  filter(interval_type == "run") |> 
  ggplot() + 
    aes(
      x = (interval_number + 1) / 2, 
      y = (time / 60) / (as.numeric(distance) / 1000)
    ) +
    geom_line(aes(group = date, color = factor(date))) +
    geom_point(aes(color = factor(date))) +
    geom_label(
      aes(
        label = paste0(distance, " m"), 
        y = 3.5
      ), 
      hjust = .5,
      stat = "unique"
    ) +
    geom_label(
      aes(label = label, y = 3.5),
      data = data.frame(interval_number = 15, label = "(all)"),
      hjust = .5,
    ) +
    geom_point(
      aes(color = factor(date)), 
      data = d_date_margins |> 
        filter(interval_type == "run") |> 
        mutate(interval_number = 15) |> 
        rename(time = total_time),
    ) +
    labs(
      x = "running interval number", 
      y = "pace [min per km] (lower: better)", 
      color = "date"
    )

d |> 
  filter(interval_type == "run") |> 
  ggplot() + 
    aes(
      x = (interval_number + 1) / 2, 
      y = (time) / (as.numeric(distance) / 200)
    ) +
    geom_line(aes(group = date, color = factor(date))) +
    geom_point(aes(color = factor(date))) +
    geom_label(
      aes(
        label = paste0(distance, " m"), 
        y = 30
      ), 
      hjust = .5,
      stat = "unique"
    ) +
    geom_label(
      aes(label = label, y = 30),
      data = data.frame(interval_number = 15, label = "(all)"),
      hjust = .5,
    ) +
    geom_point(
      aes(color = factor(date)), 
      data = d_date_margins |> 
        filter(interval_type == "run") |> 
        mutate(interval_number = 15) |> 
        rename(time = total_time),
    ) +
    labs(
      x = "running interval number", 
      y = "seconds per 200 m (lower: better)", 
      color = "date"
    )

d |> 
  filter(interval_type == "recover") |> 
  ggplot() + 
    aes(
      x = (interval_number + 1) / 2, 
      y = time
    ) +
    geom_line(aes(group = date, color = factor(date))) +
    geom_point(aes(color = factor(date))) +
    geom_label(
      aes(
        label = paste0(distance, " m"), 
        y = 45
      ), 
      hjust = .5,
      stat = "unique"
    ) +
    labs(
      x = "running interval number", 
      y = "recovery duration [s] (lower: better)", 
      color = "date",
      caption = "I used a different workout app on 2022-08-16 🤷‍♀️"
    )
```
